---
title: "FP_Stat184"
author: "Bryan Xiao, Daiwik Kashyap"
format:
  pdf:
    embed-resources: true
    toc: true
editor: visual
---

```{r}
#| label: librariesAndData
#| echo: true
#| message: false
# Load necessary libraries
library(tidytext)
library(tidyverse)
library(googlesheets4)
library(ggplot2)
library(dcData)
library(knitr)
library(tinytex)
library(stringr)
library(scales)

# Load datasets
MostPlayedDataset <- read.csv("~/GitHub/Sec4_FP_BryanXiao_DaiwikKashyap/data/data.csv",header=TRUE,row.names=)
SteamStoreDataset <- read.csv("~/GitHub/Sec4_FP_BryanXiao_DaiwikKashyap/data/steam.csv")
```

# Data Wrangling

We cleaned and merged the datasets Steam Store Games and Most Played Games of All Time, and isolated the top 200 games based on peak players.

```{r}
#| label: dataWrangling
#| echo: true
#| message: false
# Wrangle most played dataset
MostPlayedDataset$All_time.peak <- str_replace_all(MostPlayedDataset$All_time.peak, ",", "")
MostPlayedDataset$All_time.peak <- as.numeric(as.character(MostPlayedDataset$All_time.peak))

# Merge datasets and tidy
MergedData <- merge(SteamStoreDataset, MostPlayedDataset)

MergedDataTidy <- MergedData %>% 
  arrange(desc(All_time.peak), .by_group = TRUE) %>%
  select("Name", "genres", "All_time.peak", "positive_ratings", "negative_ratings") %>%
  rename(
    Genres = genres,
    All_Time_Peak = All_time.peak,
    Positive_Ratings = positive_ratings,
    Negative_Ratings = negative_ratings
  ) %>%
  slice(1:400) %>%
  separate(
    col = "Genres",
    sep = ";",
    into = c("Genre 1", "Genre 2", "Genre 3"),
    fill = "right"
  )

write.csv(
  MergedDataTidy,
  file = "Cleaned_Data.csv",
  row.names = TRUE
)

```

# Visualizations

## Top 10 Most Played Steam Games

```{r}
#| label: top10Games
#| echo: true
#| message: false
Top10Games <- MergedDataTidy %>% arrange(desc(All_Time_Peak)) %>% slice(1:10)

ggplot(Top10Games, aes(x = reorder(Name, All_Time_Peak), y = All_Time_Peak)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 10 Most Played Steam Games",
       x = "Game",
       y = "All Time Peak Players") +
  scale_y_continuous(labels = comma) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

## Count of Games by Primary Genre

```{r}
#| label: countingGameGenres
#| echo: true
#| message: false
ggplot(MergedDataTidy, aes(x = `Genre 1`)) +
  geom_bar(fill = "darkgreen") +
  coord_flip() +
  labs(title = "Count of Games by Primary Genre",
       x = "Primary Genre",
       y = "Number of Games") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
#| label: top10Genres
#| echo: true
#| message: false


ggplot(Top10Genres, aes(x = reorder(Genre, n), y = n)) +
  geom_col(fill = "darkgreen") +
  coord_flip() +
  labs(
    title = "Top 10 Most Common Genres",
    x = "Genre",
    y = "Number of Games"
  ) +
  theme_minimal()

```

## Sub-Genres in Action

Shows the sub genres of action games

```{r}
#| label: subGenreMergeData
#| echo: true
#| message: false

# Prepare a subset with only top 3 genres
TopGenres <- c("Indie", "Action", "Adventure")

# Filter and explode tags
TopSubGenres <- MergedData %>%
  filter(str_detect(genres, paste(TopGenres, collapse = "|"))) %>%
  select(Name, genres, steamspy_tags) %>%
  separate_rows(genres, sep = ";") %>%
  filter(genres %in% TopGenres) %>%
  separate_rows(steamspy_tags, sep = ";") %>%
  filter(steamspy_tags != "") %>%
  group_by(genres, steamspy_tags) %>%
  summarise(count = n(), .groups = "drop") %>%
  arrange(genres, desc(count)) %>%
  group_by(genres) %>%
  slice_max(count, n = 5)  # top 5 sub-genres for each genre

```


```{r}
#| label: topSubGeners
#| echo: true
#| message: false


ggplot(TopSubGenres, aes(x = genres, y = count, fill = steamspy_tags)) +
  geom_col() +
  coord_flip() +  # ← Flip axes
  labs(
    title = "Top 5 Sub-Genres Within Top 3 Genres",
    x = "Primary Genre",
    y = "Number of Games",
    fill = "Sub-Genre Tags"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold")
  )

```

```{r}
#| label: reviewsVsPlayerCount
#| echo: true
#| message: false
MergedDataReviews <- MergedDataTidy %>% 
  slice(1:400) %>%
  select(Name, All_Time_Peak, Positive_Ratings, Negative_Ratings) %>%
  mutate(
    PercentageOfPositiveReviews = Positive_Ratings / (Negative_Ratings + Positive_Ratings) * 100
  )
ggplot(
  data = MergedDataReviews,
  mapping = aes(
    x = All_Time_Peak,
    y = PercentageOfPositiveReviews
  )
) + geom_point(size = 3) +
  geom_smooth(method = "lm") +
  labs(
    x = "All Time Peak",
    y = "Percentage of Positive Reviews",
    title = "Percentage of Positive Reviews vs All Time Peak"
  ) + scale_x_continuous(labels = comma) +
  scale_y_continuous(labels = comma) +
  theme_bw()

```

# Conclusion

The Steam gaming reveals multiple intriguing patterns which become visible through visual analysis. Firstly,

The Top  10 Most Played Games chart shows Counter-Strike: Global Offensive and Dota 2 as top games  which sustain more than one million peak concurrent players.

The data shows competitive multiplayer games remain popular throughout time  according to this information.

The Genre Count chart reveals that Action and Indie genres make up the biggest collection  of available Steam titles. The tools which allow indie developers to develop and distribute games in these genres could  be the reason for this situation.

The Peak Players by Genre boxplot reveals that user engagement levels differ  substantially between different genres. Free to Play and Action games achieve the highest player count peaks compared to all  other genres but Nudity and Animation & Modeling genres show very low user engagement.

The data reveals  that genre selection acts as a strong indicator to determine both audience size and user participation levels.

This analysis gives us  a current view of Steam user preferences together with publishing behavior on the platform. The analysis provides important knowledge  about game development and marketing strategies as well as user conduct patterns.

Thank you for taking the time to explore our analysis!

```{r codeAppend, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```